Reciprocal Tolerance: Why States Allow Their Adversaries to Practice Espionage with Minimal Consequences

### 2020-11-08 article.tex
Conform to Security Studies Style Guidelines

\documentclass[14pt]{extarticle}
\usepackage{style}

\title{Reciprocal Tolerance: Why States Allow Their Adversaries to Practice Espionage with Minimal Consequences}
\author{Alexander W. Petros}
\date{October 2020}

@@

\begin{document}
\maketitle

@@

\section{Conclusion}
The promise of an international community devoid of secrets is, at its core, a utopian promise. Complete, universal verification of capabilities and intents would make it impossible for an adversary to maintain hidden military capabilities or furtive plans for war; theoretically, if applied to our current system of nation-states, bad actors would be rendered obsolete. While diplomacy based on perfect information is impossible right now---and likely always will be---espionage norms that reduce the diplomatic impact of intelligence gathering are one means to approach a truly open society.

In order to best demonstrate the novel theory of this paper, I focused on a uniquely evidence-rich context: two nuclear superpowers who both maintained robust intelligence services. This is, of course, not the situation in which most espionage norms live---the stakes might be lower, the capabilities different, or the overall balance of power more lopsided. For example, the relationship between the espionage services of Israel and Iran could look very different from the US-USSR espionage relationship examined here, as could the relationship between the US and present-day Russia. This paper does not claim to consider all these variations, rather, it lays a foundation for future analysis of how other variables impact the espionage norms observed.

@@

The value of an international community based on perfect information is still under contention. Despite recent worrying developments, at the time of writing no ASAT has even been used against an adversary's satellite, and the Open Skies treaty remains precariously in effect. Maintaining these norms requires formulating a positive case for their utility, and quickly. The need to present this case in a compelling manner will only become more urgent.

>>>>

\documentclass[12pt]{extarticle}
\usepackage{style}

\title{Reciprocal Tolerance: Why States Allow Their Adversaries to Practice Espionage with Minimal Consequences}
\author{Alexander W. Petros\thanks{This article would not be possible without the unwavering attention of Professor William Wohlforth, who continued to provide gentle encouragement and thoughtful criticism throughout its creation. I would also like to thank the Dartmouth Government department for its support in developing the thesis on which this article is based, and from which much of this research derives. Finally, my first and most valued editor is my father, William Petros, whose unsparing attention to detail is deeply reflected in this article, as in all things he does.}}
\date{October 2020}

@@

\begin{document}

\maketitle
\newpage
\input{abstract.tex}
\newpage

@@

\section{Conclusion}
The promise of an international community devoid of secrets is, at its core, utopian. Complete, universal verification of capabilities and intents would make it impossible for an adversary to maintain hidden military capabilities or furtive plans for war; theoretically, if applied to our current system of nation-states, bad actors would be rendered obsolete. While diplomacy based on perfect information is impossible right now---and likely always will be---espionage norms that reduce the diplomatic impact of intelligence gathering are one means to approach a truly open society.

In order to best demonstrate the novel theory of this paper, I focused on a uniquely evidence-rich context: two nuclear superpowers who both maintained robust intelligence services. This is, of course, not the situation in which most espionage norms exist---the stakes might be lower, the capabilities different, or the overall balance of power more lopsided. For example, the relationship between the espionage services of Israel and Iran could look very different from the US-USSR espionage relationship examined here, as could the relationship between the US and present-day Russia. This paper does not claim to consider all these variations, rather, it lays a foundation for future analysis of how other variables impact the espionage norms observed.

@@

The viability of international community based on perfect information is a question that remains to be settled. Despite recent worrying developments, there are still reasons to be optimistic: at the time of writing no ASAT has even been used against an adversary's satellite, and the Open Skies treaty remains precariously in effect. Maintaining these norms urgently requires formulating a compelling case for their utility.
